Optimizing future biodiversity sampling by citizen scientists

### 2019-04-25 paper/main.tex
Adding additional figure to manuscript

\subsection*{Applications of our predictions}
Providing dynamic feedback to citizen science participants has proved successful for many citizen science projects \cite{rowley2019frogid, wiggins2011conservation, xue2016avicaching}. This feedback is generally in the form of leaderboards, presenting the number of submissions or number of unique species someone has contributed \cite{wood2011ebird}. But leaderboards tend to focus on outputs, incentivizing finding rather than looking, leading to perverse outcomes related to the redundancies and gaps in citizen science data. We sought to develop an outcome-based incentive by using our fitted models to predict the expected value of a given citizen science observation, dynamically, for any given day (e.g., Fig. 3, \href{https://github.com/coreytcallaghan/optimize_citizen_science_obs/blob/master/Figures/dynamic_map.gif}{Video 2}). This approach required us to look backward first, using a model with all observations for 2018, based on statistical leverage calculated from 2010-2018, in order to look forward and predict the expected leverage for any given day. We envision a dynamic approach (\href{https://github.com/coreytcallaghan/optimize_citizen_science_obs/blob/master/Figures/dynamic_map.gif}{Video 2}) in future citizen science projects, which would ultimately guide participants to sites which \textit{should} be sampled on any given day --- or in a given week, month, or year. In this instance, leaderboards would move past numbers of species or submissions and could be derived based on a participant's cumulative value to the citizen science dataset, for biodiversity outcomes. Instead of participants preferentially chasing specific species, this approach would guide participants to the sites with the highest expected marginal value. For example, we imagine visitor centers across the world at national parks or urban greenspaces providing their visitors on any given day a localized map showing which trail someone should visit if they are interested in contributing the greatest value to that park's biodiversity knowledge, through citizen science. The global pull of ecotourism \cite{sharpley2006ecotourism} is increasing exponentially, creating the potential for people to contribute to local biodiversity knowledge in areas that are traditionally undersampled, and with this framework, the collective effort of citizen scientists can be maximized.

@@

@@

>>>>

@@

\subsection*{Applications of our predictions}
Providing dynamic feedback to citizen science participants has proved successful for many citizen science projects \cite{rowley2019frogid, wiggins2011conservation, xue2016avicaching}. This feedback is generally in the form of leaderboards, presenting the number of submissions or number of unique species someone has contributed \cite{wood2011ebird}. But leaderboards tend to focus on outputs, incentivizing finding rather than looking, leading to perverse outcomes related to the redundancies and gaps in citizen science data. We sought to develop an outcome-based incentive by using our fitted models to predict the expected value of a given citizen science observation, dynamically, for any given day (e.g., Fig. 3, Fig. 4, \href{https://github.com/coreytcallaghan/optimize_citizen_science_obs/blob/master/Figures/dynamic_map.gif}{Video 2}). This approach required us to look backward first, using a model with all observations for 2018, based on statistical leverage calculated from 2010-2018, in order to look forward and predict the expected leverage for any given day. We envision a dynamic approach (\href{https://github.com/coreytcallaghan/optimize_citizen_science_obs/blob/master/Figures/dynamic_map.gif}{Video 2}) in future citizen science projects, which would ultimately guide participants to sites which \textit{should} be sampled on any given day --- or in a given week, month, or year. In this instance, leaderboards would move past numbers of species or submissions and could be derived based on a participant's cumulative value to the citizen science dataset, for biodiversity outcomes. Instead of participants preferentially chasing specific species, this approach would guide participants to the sites with the highest expected marginal value. For example, we imagine visitor centers across the world at national parks or urban greenspaces providing their visitors on any given day a localized map showing which trail someone should visit if they are interested in contributing the greatest value to that park's biodiversity knowledge, through citizen science. The global pull of ecotourism \cite{sharpley2006ecotourism} is increasing exponentially, creating the potential for people to contribute to local biodiversity knowledge in areas that are traditionally undersampled, and with this framework, the collective effort of citizen scientists can be maximized.

@@

\begin{figure}[!hb]
\centering
\includegraphics[width=.8\linewidth]{example_grids_through_time.png}
\caption{The changes in expected marginal value for 2018 for 10 randomly chosen grid cells. Some sites' expected value remains relatively constant through time (e.g., grid 371) while others are highly variable (e.g., grid 274), and others undergo distinct step-changes (e.g., grid 405). This is a dynamic approach, showing how the expected value of a grid changes based on our parameterized model results (Fig. 3 \href{https://github.com/coreytcallaghan/optimize_citizen_science_obs/blob/master/Figures/dynamic_map.gif}{Video 2})).}
\label{fig4}
\end{figure}
